---
title: "Preparation and loading the data"
author: "B Kleinberg https://github.com/ben-aaron188"
subtitle: 'EuroCSS workshop: LTTA'
output:
  html_document: default
  pdf_document: default
---
  
### Welcome to the EuroCSS workshop on Linguistic Temporal Trajectory Analysis
  
In order to save time at the day of the workshop, we'd kindly ask you to set up your R workspace, test the functionality of your workspace with a few scripts, and download the data we will use for the paper hackathon.

This script will guide you through the necessary steps.

### Step 0: Prepare a folder that you use for this workshop.

The easiest option is to [fork](https://help.github.com/articles/fork-a-repo/) the [workshop GitHub repo](https://github.com/ben-aaron188/ltta_workshop) and clone it to your machine so you have it available and can work with full version control on it.

Alternatively you can also simply download the repository as a .zip file using the green 'Clone or download' button on the upper right part of the repository [here](https://github.com/ben-aaron188/ltta_workshop).

Once you have completed either option 1 or 2, you should have a folder called *ltta_workshop*. In that folder, you can find the general workshop information as well as a sub-directory called *"workshop_practical"*. We will use this folder in the practical part of the workshop.

Navigate to the folder `r_notebooks` and open the `workshop_preparation.Rmd` file in RStudio. To avoid clashes with paths, we recommend you run the remainder of this script by opening the `workshop_preparation.Rmd` file from the newly created folder.

### Step 1: Download the hackathon data

We have deposited the hackathon dataon the Open Science Framework at [https://osf.io/ptyv3/files/](https://osf.io/ptyv3/files/). You can download the main parent folder called `hackathon_data` as a zip and move it to the workshop folder or any other location on your machine.

We will use these datasets in the afternoon part for the paper hackathon.


### Step 2: Installing/loading required R dependencies

#### R packages from CRAN

These packages are available on CRAN and can be installed within R Studio using the command below.

```{r}
#clean your R memory
rm(list = ls())

#stringr
if (!require(stringr)){
install.packages('stringr')
} 
library(stringr)

#data.table
if (!require(data.table)){
install.packages('data.table')
} 
library(data.table)

#lexicon
if (!require(lexicon)){
install.packages('lexicon')
} 
library(lexicon)

#syuzhet
if (!require(syuzhet)){
install.packages('syuzhet')
} 
library(syuzhet)
```


#### R dependencies from GitHub/source

The *naive context sentiment* code is "live" available on [GitHub](https://github.com/ben-aaron188/naive_context_sentiment). 
- You can [fork and clone this repository to your machine](https://help.github.com/articles/cloning-a-repository/)
- or, alternatively, you can load the dependency from the main directory of the ltta_workshop folder using this command:

```{r}
#set you working directory to the folder for the remainder of this intro script
source('../workshop_practical/r_deps/naive_context_sentiment/ncs.R')
#using this method ensures you have the NCS dependency. It does not update to any commits made in the future (use the GitHub dependency for this for future use)
```

The *txt_df_from_dir.R* function loads .txt files from a (nested) directory into an R dataframe. This makes the text data (e.g., transcripts) useable for later analysis. Code is "live" available on [GitHub](https://github.com/ben-aaron188/r_helper_functions/blob/master/txt_df_from_dir.R) and in the local source. 

```{r}
source('../workshop_practical/r_deps/txt_df_from_dir.R')
```

### Step 3: Loading the workshop data into R

*Important*: The following steps assume that you are in the `./ltta_workshop/workshop_practical` folder.

In that folder, you see three sub-directories:  `r_deps`, `r_scripts`, `sample data`.

- `r_deps` contains the local versions of the dependencies needed for this workshop.
- `r_scripts` contains the scripts used for the feature extraction of the hackathon data.
- `sample data` contains data we will use to walk-through LTTA in the morning practical.

The following steps guide you through loading and testing these datasets.

#### Load the "YouTube vlogger" sample data

This dataset consists of all YouTube vlog transcripts from [Casey Neistat](https://www.youtube.com/user/caseyneistat) until April 2018 (adopted from [a recent paper using the LTTA method](https://arxiv.org/ftp/arxiv/papers/1808/1808.09722.pdf).

This script will guide you through the full, minimal LTTA workflow for a set of ten vlog transcripts.

1. Load raw YouTube transcript data from individual .txt files to an R dataframe

```{r}
ten_vlogs_casey_neistat = txt_df_from_dir(dirpath = '../workshop_practical/sample_data/caseyneistat_small'
, recursive = F
, include_processed = F
, to_lower = F)
#head(ten_vlogs_casey_neistat)
```


2. Extract dynamic sentiment shapes (= the core of LTTA on the intra-textual level)

```{r}
#this code extracts the sentiment shapes for each vlog transcript
#Note: this might take a few seconds
sentiment_shapes = ncs_full(txt_input_col = ten_vlogs_casey_neistat$text
, txt_id_col = ten_vlogs_casey_neistat$id
, low_pass_filter_size = 5
, transform_values = T
, normalize_values = F
, min_tokens = 10 #minimum length required to process a vlog
, cluster_lower = 3 #window size before sentiment
, cluster_upper = 3 #window size after sentiment
)
```

```{r}
View(sentiment_shapes)
#Note that each transcript is represented in one column with 100 rows (= standardized narrative time)

#set names to corresponding file name
names(sentiment_shapes) = ten_vlogs_casey_neistat$Filename
```


3. Analyse & visualise the sentiment shapes

```{r echo=TRUE}
#Show the sentiment shape of the vlog named '3.txt'

plot(sentiment_shapes$`3.txt`
, type='h'
, ylim = c(-1.25, 1.25)
, main = 'Shape of file 3.txt'
, ylab = 'Sentiment'
, xlab = 'Standardized narrative time')
```

```{r}
#Some descriptives: proportion of sentiment above 0

prop.table(table(sentiment_shapes$`3.txt` > 0))
```


#### Test load the "hackathon" data

The hackathon data includes two datasets: a media_data dataset of transcripts of 'left' and 'right' news channels, and a dataset of transcripts from YouTube's "Creators for Change" with matched "normal" vloggers.

Details on both datasets will be presented in the workshop on Wednesday.

```{r}
#MEDIA DATA
#load the RData file of the sampled media data data
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/main_data/eurocss_media_data_sampled.RData')

View(head(dt.sampled_balanced))
```

```{r}
#CREATORS FOR CHANGE DATA
#load the RData file of the sampled media data data
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/main_data/eurocss_cfc_data_full.RData')

View(head(dt.data))
```


We have extracted some features from the transcripts already:
  - unigrams
- bigrams
- trigrams
- LIWC variables (info [here](https://repositories.lib.utexas.edu/bitstream/handle/2152/31333/LIWC2015_LanguageManual.pdf))
- POS proportions (using the [qdap package](https://rdrr.io/cran/qdap/man/pos.html))
- sentiment shapes (as done above)

You can load these as follows:
  
  ```{r}
#load some features
##"static" features: unigrams
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/features/eurocss_cfc_data_unigrams.RData')
df.tfidf_ngrams_1[1:10, 100:120]
```

```{r}
#LIWC output
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/features/eurocss_media_data_liwc.RData')
dt.data_liwc[1:10, 50:60]
```

```{r}
##POS
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/features/eurocss_media_data_pos.RData')
dt.data_pos[1:10, 20:30]
```

```{r}
##sentiment shapes
load('PATH_TO_YOUR_DOWNLOADED_FOLDER/hackathon_data/features/eurocss_cfc_sentiment.RData')
head(dt.data_sentiment)
```

### End

If all of the above steps worked, you are set to start with the workshop and practical.

-------------------
  
  